---
title: "EOG and EOC data for NC"
editor: visual
date: last-modified
date-format: "[Last modified on] MMMM DD, YYYY HH:mm:ss zzz"
format: 
  html:
    theme: cosmo  # zephyr pulse, sketchy, default, cosmo, vapor etc
author:
  - name: Emma McLaughlin
    email: mclaughlinea@appstate.edu
code-fold: false
---

## Understanding the data

This data comes from the "2023-24 North Carolina School Assessments and Other Indicators Workbook". This data set contained many different pages on test scores and indicators from elementary school to high school, but I chose to focus on EOC and EOG scores among NC schools. First I read in the data, specifically the sheet of data I needed.

```{r}
#| warning: false
#| message: false
library(readxl)
testdata_eocg <- read_excel("Workbook5.xlsx")
```

Then I downloaded the packages I needed to visualiza the data.

```{r}
#| warning: false
#| message: false
library(ggplot2)
library(dplyr)
```

Here I changed the names of the columns and took out the first two rows of data because they were sub headers.

```{r}
#| warning: false
#| message: false
testdata_eocg <- testdata_eocg %>% 
  rename("District"= "Back to Introduction",
         "School_Code" = "...2",
         "School_Name" = "...3",
         "Board_region"= "...4",
         "Grade_span" = "...5",
         "Subgroup" = "...6",
         "Subject" = "...7",
         "Not_Proficient" = "...8",
         "Perc_Lev3" = "...9",
         "Perc_Lev4" = "...10",
         "Perc_Lev5" = "...11",
         "Perc_ab3" = "...12",
         "Perc_ab4" = "...13")

```

```{r}
#| warning: false
#| message: false
testdata_eocg <- testdata_eocg[-c(1,2), ]
```

Here I made some of the factors numeric in order to simplify the visualization.

```{r}
#| warning: false
#| message: false
testdata_eocg$Perc_ab3 <- as.numeric(testdata_eocg$Perc_ab3)
testdata_eocg$School_Code <- as.numeric(testdata_eocg$School_Code)
testdata_eocg$Not_Proficient <- as.numeric(testdata_eocg$Not_Proficient)
testdata_eocg$Perc_Lev3 <- as.numeric(testdata_eocg$Perc_Lev3)
testdata_eocg$Perc_Lev4 <- as.numeric(testdata_eocg$Perc_Lev4)
```

## Graphs of test score by school

```{r}
#| fig-cap: "This graph shows the percentage of students that scored a 3 or higher on their Reading Grade 3 EOG"
#| warning: false
testdata_eocg %>%
  filter(!(School_Code == "NC")) %>% 
  filter(Subgroup == "All Students", Subject == "Reading Grade 3") %>% 
  ggplot(aes(Perc_ab3)) +
  geom_bar(fill = "lightblue")  + 
  theme_bw() + 
  labs(x="Percent of students scoring above a 3")
```

Based on this graph, it looks like many schools have 50 % of students scoring above a 3 and most of the data is concentrated between 38%-63%.

```{r}
#| fig-cap: "This graph shows the percentage of students that scored a 3 or higher on their Math Grade 3 EOG"
#| warning: false
#| message: false
testdata_eocg %>% 
  filter(!(School_Code == "NC")) %>% 
  filter(Subgroup == "All Students", Subject == "Math Grade 3")%>%
  ggplot(aes(Perc_ab3)) + 
  geom_bar(fill="lightpink") + 
  theme_bw() + 
  labs(x="Percent of students scoring above a 3")
```

This graph shows most schools having 50% - 75% of their students scoring above a 3.

```{r}
#| warning: false
#| message: false
#| fig-cap: "Comparing percent of students scoring above a 3 in 3rd Grade Math and Reading"
testdata_eocg %>% 
  filter(!(School_Code == "NC")) %>% 
  filter(Subgroup == "All Students", Subject == "Math Grade 3"| Subject == "Reading Grade 3") %>% 
  ggplot(aes(Perc_ab3)) + geom_histogram(bins=100) +
  facet_wrap(vars(Subject))+
  theme_bw() + 
  labs(x="Percent of students scoring above a 3")
```

Through comparison we see that the test scores between subjects do not differ greatly, but percentages for Math tend to be a bit better.

```{r}
#| warning: false
#| message: false
#| fig-cap: "Comparing percent of students scoring above a 3 in 5th Grade Math and Reading"
testdata_eocg %>% 
  filter(!(School_Code == "NC")) %>% 
  filter(Subgroup == "All Students", Subject == "Math Grade 5"| Subject == "Reading Grade 5") %>% 
  ggplot(aes(Perc_ab3)) + geom_histogram(bins=100) +
  facet_wrap(vars(Subject))+
  theme_bw() + 
  labs(x="Percent of students scoring above a 3")
```

The graphs for 5th grade testing look very similar to the graphs from 3rd grade testing. There does not appear to be many differences.

```{r}
#| warning: false
#| message: false
#| fig-cap: "Seeing if there is any relationship between region and proficiency on exams"
testdata_eocg %>% 
  filter(!(School_Code == "NC")) %>% 
  filter(Subgroup == "All Students", Subject == "Reading Grade 3") %>% 
  ggplot(aes(Perc_ab3, School_Name, color = Board_region)) + geom_point() +
  theme(axis.text.y = element_blank()) + 
  labs(x = "Percent of students scoring above a 3", y= "Schools", color = "Region")
```

There is not a clear pattern between the percent of students scoring above a 3 and the region where their school is.

```{r}
#| warning: false
#| message: false
#| fig-cap: "Attempting again to find relation between proficiency and region"
testdata_eocg %>% 
  filter(!(School_Code == "NC")) %>% 
  filter(Subgroup == "All Students", Subject == "Math Grade 3-8") %>% 
  ggplot(aes(Perc_Lev3, Perc_Lev4, color = Board_region)) + geom_point() + geom_jitter() + xlim(0,40) + labs(x="Percent Scoring 3", y="Percent Scoring 4", color = "Region")
```

Even when `Schools` is no longer a variable there are no distinguishable patterns between test score and region. 

## Graphs of overall NC test data

```{r}
#| warning: false
#| message: false
#| fig-cap: "Proficiency based on subject in NC"
testdata_eocg %>% 
  filter(District == "State of North Carolina") %>% 
  ggplot(aes(Not_Proficient, Subject)) +
 geom_boxplot(fill= "purple") + 
  labs(x="Percent of students not proficient in the subject", y="Subject") + 
  theme_bw()
```

The `Subject` with the best proficiency appears to be Science . Reading and Math subjects appear to be similar to each other in terms of proficiency.

```{r}
#| warning: false
#| message: false
#| fig-cap: "Proficiency by subgroups"
testdata_eocg %>% 
  filter(!(District == "State of North Carolina")) %>% 
  ggplot(aes(Not_Proficient)) + 
  geom_histogram() + 
  facet_wrap(vars(Subgroup)) +
  theme_bw() +
  labs(x="Percent of students not proficient")
```

Groups like Homeless students, English Learners, and Disabled students appear to be less proficient in terms of testing. There also seems to be very little data on Migrant students, students in Foster Care, and Native students.
